---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
Hemant Bhambhu
========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

library(ggplot2)
library(dplyr)
library(corrplot)
library(psych)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
red_wine=read.csv('wineQualityReds.csv')
```

Information about our dataset-
This tidy data set contains 1,599 red wines with 11 variables on the chemical 
properties of the wine. At least 3 wine experts rated the quality of each wine,
providing a rating between 0 (very bad) and 10 (very excellent).
Description of attributes:

   1 - fixed acidity: most acids involved with wine or fixed or nonvolatile 
   (do not evaporate readily)

   2 - volatile acidity: the amount of acetic acid in wine, which at too high
   of levels can lead to an unpleasant, vinegar taste

   3 - citric acid: found in small quantities, citric acid can add 
   'freshness' and flavor to wines

   4 - residual sugar: the amount of sugar remaining after fermentation 
   stops, it's rare to find wines with less than 1 gram/liter and wines 
   with greater than 45 grams/liter are considered sweet

   5 - chlorides: the amount of salt in the wine

   6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between
   molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial
   growth and the oxidation of wine

   7 - total sulfur dioxide: amount of free and bound forms of S02; in low
   concentrations, SO2 is mostly undetectable in wine, but at free SO2
   concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

   8 - density: the density of water is close to that of water depending
   on the percent alcohol and sugar content

   9 - pH: describes how acidic or basic a wine is on a scale from 
   0 (very acidic) to 14 (very basic); most wines are between 3-4 
   on the pH scale

   10 - sulphates: a wine additive which can contribute to sulfur dioxide
   gas (S02) levels, wich acts as an antimicrobial and antioxidant

   11 - alcohol: the percent alcohol content of the wine


# Univariate Plots Section


```{r echo=FALSE, Univariate_Plots}

str(red_wine)

```
```{r}
M <- cor(red_wine)
corrplot(M, method="circle",type = 'lower')
```



observation for quality-
positive correlation - fixed acidity,citric acid, 
residual sugar,alcohol,sulphates

negative correlation-volatile acidity,chloride,
free sulphur dioxide, total sulfur dioxide,density,pH

it's highly corelated to alcohol and volatile acidity.

```{r echo=FALSE}


multi.hist(select(red_wine,
            c(fixed.acidity,volatile.acidity,citric.acid,residual.sugar)))
```
fixed.acidity has near normal distribution,
volatile.acidity has near normal distribution,
citric.acid has near uniform distribution
residual.sugar is not clear so we draw it again with reduced scale

```{r echo=FALSE , message=FALSE, warning=FALSE}
ggplot(aes(x=residual.sugar),data = red_wine)+
  geom_histogram(binwidth=.1)+
  coord_cartesian(xlim = c(0.8,4))+
  scale_x_continuous(breaks = seq(0,4,.25))
```
yes sugar is also fairly normally distributed.

```{r echo=FALSE}
summary(select(red_wine,
            c(fixed.acidity,volatile.acidity,citric.acid,residual.sugar)))
```
```{r echo=FALSE}
red_wine$quality_text=as.factor(cut(as.numeric(red_wine$quality),
c(2,4,6,8),labels = c('bad','average','good')))
red_wine$quality_text=as.factor(red_wine$quality_text)

```

```{r echo=FALSE}
red_wine$residual.sugar_content=
  as.factor(cut(as.numeric(red_wine$residual.sugar),
c(0,1.9,2.6,16),labels = c('low','average','high')))
red_wine$residual.sugar_content=as.factor(red_wine$residual.sugar_content)

```

```{r echo=FALSE}
red_wine$fixed.acidity_content=
  as.factor(cut(as.numeric(red_wine$volatile.acidity),
c(0,7.1,9.2,16),labels = c('low','average','high')))
red_wine$fixed.acidity_content=as.factor(red_wine$fixed.acidity_content)

```

```{r echo=FALSE}
multi.hist(select(red_wine,
          c(chlorides,free.sulfur.dioxide,total.sulfur.dioxide,sulphates)))
```
```{r echo=FALSE}
summary(select(red_wine,
            c(chlorides,free.sulfur.dioxide,total.sulfur.dioxide,sulphates)))
```
sulphur is comparitively higher when considering only numerical value let's 
plot this on logarithmic scale. 

```{r echo=FALSE}
red_wine$free.sulfur.dioxide_content=
  as.factor(cut(as.numeric(red_wine$free.sulfur.dioxide),
c(0,7,21,73),labels = c('low','average','high')))
red_wine$free.sulfur.dioxide_content=
  as.factor(red_wine$free.sulfur.dioxide_content)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=total.sulfur.dioxide),data = red_wine)+
  geom_histogram(binwidth=.1)+
  scale_x_log10()
```
total.sulfur.dioxide distribution seemed to be right skewed is actually 
normally when viewed in logarithmic scale.

distribution of chlorides was not clear so let's plot it with reduced scale.
```{r echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=chlorides),data = red_wine)+
  geom_histogram()+
  coord_cartesian(xlim = c(0,.7))+
  scale_x_continuous(breaks = seq(0,.7,.05))
```


```{r echo=FALSE}
multi.hist(select(red_wine,c(pH,alcohol)))
```
```{r echo=FALSE}
summary(select(red_wine,
            c(pH,alcohol)))
```

Almost all of our feature variabls area nearly normally distributed. 
citric acid has bimodal distribution.

```{r echo=FALSE}
ggplot(aes(x=citric.acid,fill=quality_text),data = red_wine)+
  geom_density(alpha=.3)
```
```{r echo=FALSE}
red_wine$alcohol_content=as.factor(cut(as.numeric(red_wine$alcohol),
c(0,9.5,11.1,15),labels = c('low','average','high')))
red_wine$alcohol_content=as.factor(red_wine$alcohol_content)

```

# Univariate Analysis



### What is the structure of your dataset?
There are 1599 red wine in this data set with 12 features 
(fixed.acidity, volatile.acidity, citric.acid, residual.sugar, chlorides, 
free.sulfur.dioxide, total.sulfur.dioxide, density,pH,sulphates,alcohol,quality)
Most of Features are normally distributed. Except Sulfur.dioxide
(both Free and Total) and alcohol other features are narrowly distributed.

### What is/are the main feature(s) of interest in your dataset?
Quality has high corelation with alcohol and fixed acidity. these are main 
feature of interest. 


### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

i would also like to explore relationship with citric acid
and sulphur content.i would like to explore features with correlation greater 
than .2 in magnitude.

### Did you create any new variables from existing variables in the dataset?

yes i have converted some features into catogorical with mainly three catogories
low,average and high. Since our dataset in not big it will help in highlighting 
main distinguishing features.

### Of the features you investigated, were there any unusual distributions? 

No, most features follow normal distribution but there are some outliers in
some distributions like in sulphur dioxide.



# Bivariate Plots Section




```{r echo=FALSE}
ggplot(data = red_wine,aes(x=pH,y=fixed.acidity))+
  geom_point()+
  geom_smooth(method = 'lm')

```
```{r echo=FALSE}
ggplot(data = red_wine,aes(x=density,y=fixed.acidity))+
  geom_point()+
  geom_smooth(method = 'lm')

```

```{r echo=FALSE}
ggplot(data = red_wine,aes(x=citric.acid,y=fixed.acidity))+
  geom_point()+
  geom_smooth(method = 'lm')

```

```{r echo=FALSE}
ggplot(data = red_wine,aes(x=citric.acid,y=volatile.acidity))+
  geom_point()+
  geom_smooth(method = 'lm')

```

```{r echo=FALSE}
ggplot(data = red_wine,aes(x=density,y=alcohol))+
  geom_point()+
  geom_smooth(method = 'lm')

```

```{r echo=FALSE}
ggplot(data = red_wine,aes(x=free.sulfur.dioxide,y=total.sulfur.dioxide))+
  geom_point()+
  geom_smooth(method = 'lm')

```
```{r echo=FALSE}
ggplot(data = red_wine,aes(x=volatile.acidity,y=pH))+
  geom_point()+
  geom_smooth(method = 'lm')

```

```{r echo=FALSE}
ggplot(data = red_wine,aes(x=quality_text,y=volatile.acidity))+
  geom_boxplot()

```
```{r echo=FALSE}
ggplot(data = red_wine,aes(x=quality_text,y=alcohol))+
  geom_boxplot()

```
```{r echo=FALSE}
ggplot(data = red_wine,aes(x=quality_text,y=citric.acid))+
  geom_boxplot()+
  geom_smooth(method = 'lm')+
  coord_cartesian(ylim = c(0,.75))

```

```{r echo=FALSE}
ggplot(data = red_wine,aes(x=quality_text,y=sulphates))+
  geom_boxplot()+
  geom_smooth(method = 'lm')+
  coord_cartesian(ylim = c(0.25,1))

```



# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Quality is positively associated with Sulphates,citric acid and alcohol content.
Quality is negatively associated with volatile acidity.
fixed acidity is highly corelated with ph,citric acid and density.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

There is positive corelation between volatile acidity and pH which is pretty
surprising.one speculation is that because volatile acidity is negatively 
corelated with citric acid so increases in volatile acidity makes solution less
acidic.

### What was the strongest relationship you found?
Posotive-
citric acid and fixed acidity
density and fixed acidity
quality and alcohol
Negative-
ph and fixed acidity
citric acid and volatile acidity
alcohol and density
quality and volatile acidity

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

```{r echo=FALSE}
ggplot(data = red_wine,aes(x=alcohol,y=density))+
  geom_point(aes(color=quality_text))

```
```{r echo=FALSE}
ggplot(data = red_wine,aes(x=volatile.acidity,y=citric.acid))+
  geom_point(aes(color=quality_text))

```
```{r echo=FALSE}
ggplot(data = red_wine,aes(x=pH,y=citric.acid))+
  geom_point(aes(color=quality_text),alpha=.4)

```
```{r echo=FALSE}
ggplot(data = red_wine,aes(x=alcohol_content,y=volatile.acidity))+
  geom_point(aes(color=quality_text),alpha=.3)+
  facet_grid(~quality_text)

```


```{r echo=FALSE}
ggplot(data = red_wine,aes(x=pH,y=citric.acid))+
  geom_bin2d(aes(fill=quality_text))+
  facet_grid(~residual.sugar_content)

```

```{r echo=FALSE}
ggplot(data = red_wine,aes(x=fixed.acidity,y=density))+
  geom_point(aes(color=quality_text))+
  facet_wrap(~alcohol_content)

```


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
Low volatile and high citric acid is recipe for good wine.

### Were there any interesting or surprising interactions between features?
No
------

# Final Plots and Summary



### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x=alcohol),data = red_wine)+
  ggtitle('Quality dependence on Alcohol content')+
 geom_density(aes(color=quality_text,fill=quality_text,alpha=.01))+
  scale_x_continuous(name = 'Alcohol content in wine',breaks = seq(8,16,.5))+
  scale_y_continuous(name = 'Density')
```

### Description One
Quality is most corelated with alcohol content.we can clearly see that good 
quality wines has significantly higher amount of alcohol present

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(data = red_wine,aes(x=quality_text,y=volatile.acidity))+
  geom_boxplot()+
  ggtitle('Quality dependence on Volatile acidity')+
  scale_x_discrete(name = 'Quality of Wine')+
  scale_y_continuous(name = 'Volatile Acidity',breaks = seq(0,1.6,.2))
```

### Description Two
volatile acidity which at too high of levels can lead to an unpleasant, vinegar
taste is clearly validated through this graph.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(data = red_wine,aes(x=volatile.acidity,y=citric.acid))+
  geom_point(aes(color=quality_text))+
  ggtitle('Quality dependence on Volatile acidity and citric acid')+
  scale_x_continuous(breaks = seq(0,1.6,.2), name = 'Volatile Acidity')+
  scale_y_continuous(name = 'Citric Acid Content',breaks = seq(0,1,.1))
```

### Description Three
Low volatile and high citric acid is recipe for good wine.
------

# Reflection

Our data set contains 1,599 red wines with 11 variables on the chemical 
properties of the wine.
Our Dataset had no missing values. we have six quality classes. About 82% wines
were of quality 5,6.so when plotting it was diffucult to see other quality
representation by different features. so i decided to group quality into 3
classes bad,average and good. these changes helped in visualizing more clearly.

Based on our intution we checked dependencies of quality on different features.
most of the features were normally distributed except citric acid which had 
bimodal distribution but on logarithmic scale it was also near normal 
distribution.Based on corelation about half of the feature seem to be 
positively associated and other half negatively.

based on our intution we checked quality dependence on different features.
we confirmed that the amount of acetic acid in wine, which at too high of levels
can lead to an unpleasant, vinegar taste. wines are acidic in nature. 
Quality is most corelated with alcohol content.we can clearly see that good 
quality wines has significantly higher amount of alcohol present
Low volatile and high citric acid is recipe for good wine.





